---
title: "COHA Dispersal Ridgeline Analysis"
format: 
  html:
    code-fold: true
    toc: true
    theme: cosmo
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(ggplot2)

# Load data
data <- read.csv("data/data.csv")

# Source the plot function
source("plot_function.R")
```

## Overview

This report presents 20 ridgeline plots analyzing COHA dispersal masses across 6-year periods. The first 10 plots use a compact scale (0.85), while the second 10 use an expanded scale (2.25). Each plot uses a unique color palette pair.

## Summary

This analysis presents COHA dispersal mass distributions across 6-year generational periods using ridgeline plots. Each plot includes:

- **Ridgeline densities**: Distribution of Unknown-dispersed bird masses
- **Colored triangles**: Mean mass of Unknown-dispersed birds per period (at baseline)
- **Colored dots**: Mean mass of Wisconsin-dispersed birds per period
- **Solid colored lines**: Visual guide for Unknown dispersal means (matches ridge color)
- **Dashed black lines**: Visual guide for Wisconsin dispersal means

### Analysis Structure

- **Compact plots (1-10)**: Scale = 0.85, line height = 0.85, ideal for dense overviews
- **Expanded plots (11-20)**: Scale = 2.25, line height = 1.0, ideal for detailed inspection
- **Palette pairs**: Each plot uses unique color combinations to demonstrate visual alternatives

## Compact Scale Plots (Scale = 0.85)

### Plot 1: Plasma & Inferno

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "plasma", color_palette = "inferno", 
                      palette_type = "viridis")
```

### Plot 2: Viridis & Magma

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "viridis", color_palette = "magma", 
                      palette_type = "viridis")
```

### Plot 3: Magma & Plasma

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "magma", color_palette = "plasma", 
                      palette_type = "viridis")
```

### Plot 4: Inferno & Viridis

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "inferno", color_palette = "viridis", 
                      palette_type = "viridis")
```

### Plot 5: Cividis & Rocket

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "cividis", color_palette = "rocket", 
                      palette_type = "viridis")
```

### Plot 6: Rocket & Mako

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "rocket", color_palette = "mako", 
                      palette_type = "viridis")
```

### Plot 7: Mako & Turbo

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "mako", color_palette = "turbo", 
                      palette_type = "viridis")
```

### Plot 8: Turbo & Cividis

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "turbo", color_palette = "cividis", 
                      palette_type = "viridis")
```

### Plot 9: Set2 & Dark2

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "Set2", color_palette = "Dark2", 
                      palette_type = "brewer")
```

### Plot 10: Spectral & RdYlBu

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 0.85, line_height = 0.85, 
                      fill_palette = "Spectral", color_palette = "RdYlBu", 
                      palette_type = "brewer")
```

## Expanded Scale Plots (Scale = 2.25)

### Plot 11: Plasma & Inferno

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "plasma", color_palette = "inferno", 
                      palette_type = "viridis")
```

### Plot 12: Viridis & Magma

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "viridis", color_palette = "magma", 
                      palette_type = "viridis")
```

### Plot 13: Magma & Plasma

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "magma", color_palette = "plasma", 
                      palette_type = "viridis")
```

### Plot 14: Inferno & Viridis

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "inferno", color_palette = "viridis", 
                      palette_type = "viridis")
```

### Plot 15: Cividis & Rocket

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "cividis", color_palette = "rocket", 
                      palette_type = "viridis")
```

### Plot 16: Rocket & Mako

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "rocket", color_palette = "mako", 
                      palette_type = "viridis")
```

### Plot 17: Mako & Turbo

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "mako", color_palette = "turbo", 
                      palette_type = "viridis")
```

### Plot 18: Turbo & Cividis

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "turbo", color_palette = "cividis", 
                      palette_type = "viridis")
```

### Plot 19: Set2 & Dark2

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "Set2", color_palette = "Dark2", 
                      palette_type = "brewer")
```

### Plot 20: Spectral & RdYlBu

```{r}
#| fig-width: 10
#| fig-height: 7
create_ridgeline_plot(data, scale_value = 2.25, line_height = 1, 
                      fill_palette = "Spectral", color_palette = "RdYlBu", 
                      palette_type = "brewer")
```
